---
title: "Take-home_Ex01"
format: 
  html:
    code-fold: true
    code-summary: "Show code"
execute: 
  eval: true
  echo: true
  warning: false
date: "`r Sys.Date()`"
---

WIP \# Objective We will be applying appropriate spatial point patterns analysis methods learned in class to discover the geographical and spatio-temporal distribution of Grab hailing services locations in Singapore.

# Getting Started

## Loading R packages

The R packages that we will be using in this exercise are as follows: - arrow: For reading parquet files (Grab-Posisi Dataset) - lubridate: To handle the date formatting - sf: Import, manage and process vector-based geospatial data in R. - spatstat: Wide range of useful functions for point pattern analysis and derive kernel density estimation (KDE) layer. - tidyverse - spNetwork - tmap - classInt - viridis

```{r}
#pacman::p_load(arrow, lubridate, sf, tidyverse, spNetwork, tmap, classInt, viridis)
```

## Importing the datasets

The datasets that we will be using are: - Grab-Posisi (Aspatial) - Road data set from OpenStreetMap - Master Plan 2019 Subzone Boundary (No Sea) from Data.gov.sg

# Grab-Posisi Dataset

```{r}
#df <- read_parquet("data/aspatial/part-00000.snappy.parquet")

#df$pingtimestamp <- as_datetime(df$pingtimestamp)
```

```{r}
#append
#df1 <- rbind(df,df)
```

# OpenStreetMap Road Dataset

```{r}
#sg_road <- st_read(dsn = "data/geospatial", layer = "gis_osm_roads_free_1") %>% st_transform(crs = 3414)
#st_crs(sg_road)
```

```{r}
#st_crs(sg_road)
```

```{r}
#plot(head(sg_road))
```

# Master Plan 2019 Subzone Boundary (No Sea) Dataset

```{r}
#mpsz2019 <- st_read("data/geospatial/MasterPlan2019SubzoneBoundaryNoSeaKML.kml")

#st_crs(mpsz2019)
```

```{r}
#glimpse(mpsz2019)
```

```{r}
# st_intersects(mpsz2019, sg_road)
```
